---
title: "Practical Machine Learning Course Project"
author: ""
date: ""
output: html_document
---

##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants are used. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of the project is to predict the manner in which they did the exercise, which is the "classe" variable in the training set

##Data


The data for this project comes from this source: http://groupware.les.inf.puc-rio.br/har. We download the training and testing datasets from

*https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

*https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The two data sets were then imported into R for model building.

```{r}
training<-read.csv("C://Users/Dehui/Documents/datascience/PracticalMachineLearning/pml-training.csv")
testing<-read.csv("C://Users/Dehui/Documents/datascience/PracticalMachineLearning/pml-testing.csv")
```
The training data has 19622 observations and 160 variables. The "classe" variable has five levels A,B,C,D,E.

```{r echo=FALSE, results='hide',message=FALSE}
dim(training)    
table(training$classe)
```

### Preprocessing

We separate our training data into a training set and a validation set so that we can validate our model.

```{r}
library(caret)
set.seed(9018)
trainset <- createDataPartition(training$classe, p = 0.7, list = FALSE)
Training <- training[trainset, ]
Validation <- training[-trainset, ]
```

Cleaning up near zero variance features, columns with missing values and descriptive fields.

```{r}
nzvcol <- nearZeroVar(Training)
Training <- Training[, -nzvcol]
``` 

Removing variables with 40% more missing values and descriptive variabels.

```{r}
cntobv<- sapply(Training, function(x) {
    sum(!(is.na(x) | x == ""))
})
rmnames1<- names(cntobv[cntobv< 0.6 * length(Training$classe)])
rmnames2 <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")

Training <- Training[, !names(Training) %in% c(rmnames1,rmnames2)]
```

### Model Building

We usse random forest algorithm as in the randomForest package - Breiman and Cutler's Random Forests for Classification and Regression.

```{r echo=FALSE}
library(randomForest)
rfModel <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 8)
```


Training set accuracy

```{r}
ptraining <- predict(rfModel, Training)
print(confusionMatrix(ptraining, Training$classe))
```

The model appears perform well on the training set with an acuracy of 100%.

###Model Validation

Let us now see how our model performs on the cross validation set.

```{r}
pvalidation <- predict(rfModel, Validation)
print(confusionMatrix(pvalidation, Validation$classe))
```

The cross validation accuracy is 99.6% and the out-of-sample error is therefore 0.4%. The model performs well.

###Testing Set Prediction

Applying the model to the testing set

```{r}
ptesting <- predict(rfModel, testing)
ptesting
```

Generating output files for submission

```{r}
answers <- as.vector(ptesting)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)

summary(cars)
```

